---
title: "02-prep-coralcover-gridded.Rmd"
author: "Whitney Friedman"
date: "5/16/2022"
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango  

---
# Combine datasets for use in coral counterfactual
NEW: updated from 03-combine-datasets.Rmd on 5/16/2022. 
Averaging coral cover to 5km grids

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r message=FALSE, warning=FALSE}
library(here)
source(here("globalprep","analysis","coral-counterfactual", "00-load-cc-libraries.R"))
source(here("globalprep","analysis","coral-counterfactual", "00-load-cc-functions.R"))

```


# Load & Extract Data

# 1. Response - Coral cover
## 1a. MERMAID
- Source: MERMAID
- Variables: country, site, pct_cover_hard_coral_combined, depth_avg, aca_geomorphic, management_rules
- ACA alternative: Extract from ACA directly - [see tutorial here]
(https://github.com/CoralMapping/tutorials/blob/main/WMS/eez_percent_distribution_tutorial.ipynb)
```{r}
# Load Mermaid data
source(here("globalprep","analysis","coral-counterfactual","01-format-mermaid-public.R"))

## Keep only cols used for this analysis
## Keep only data collected >= 2010

dat_mer <- cc_mermaid %>% 
  select(country, site, latitude, longitude, 
         sample_date, year,
         pct_cover_hard_coral_combined, 
         depth_avg, aca_geomorphic, management_rules, 
         db, id, cc_id) %>% 
  filter(year >= 2010) %>% 
  mutate(mgmt_simple = case_when(management_rules == "no take" ~ "no take",
                                 management_rules == "open access" ~ "open access",
                                 str_detect(management_rules, "restriction") ~ "restricted", 
                                 str_detect(management_rules, "periodic") ~ "restricted")) %>%
  mutate(aca_geomorphic = na_if(aca_geomorphic, "NA"))
  
dat_mer

# Some data checks ...
dat_mer %>% tabyl(management_rules)
dat_mer %>% tabyl(mgmt_simple)
dat_mer %>% tabyl(aca_geomorphic)

# n surveys
dat_mer %>% tabyl(country, mgmt_simple) %>% adorn_totals()

# n sites
dat_mer %>% 
  distinct(country, site, .keep_all = T) %>%
  tabyl(country, mgmt_simple) %>% 
  adorn_totals()

```

###  Update as needed
- For sites surveyed over multiple years; AVERAGE
```{r}

dat_mer1 <- dat_mer

# 1. Convert all FJ "open access" sites to "restricted" 
dat_mer1 <- dat_mer1 %>% 
  mutate(mgmt_simple = 
           case_when((country == "Fiji" & mgmt_simple == "open access") ~ "restricted",
                     TRUE ~ mgmt_simple)) %>% 
  rename(method = db) %>% 
  mutate(db = "mermaid") %>% 
  rename(pct_hardcoral = pct_cover_hard_coral_combined)


# 2. Read in updated mgmt category file from Emily (1/10/2022)
dat_mer_fix <- read_csv(here("globalprep","analysis", "coral-counterfactual","intermediate-files",
                              "cc-sites-multiple-mgmt-categories-esd.csv")) %>% 
   select(country,site,longitude,latitude, mgmt_use) %>% 
   mutate(mgmt_replace = 1)

# Replace values as indicated, final updates: 
dat_mer1 <- dat_mer1 %>% 
  left_join(dat_mer_fix, by = c("country","site","longitude","latitude")) %>% 
   mutate(mgmt_simple = case_when(mgmt_replace == 1 ~ mgmt_use, 
                                  TRUE ~ mgmt_simple)) %>% 
   select(-mgmt_replace, mgmt_use)

#1929 surveys
dat_mer1 %>% 
  tabyl(mgmt_simple) %>% 
  adorn_totals()
```


## 1b. Indo-Pacific / NEE data
- Import cc data & co-variates from Darling et al 2019
- Align with counterfactual model data (ccmod_df)
- Variables: 
-- Coral cover
-- country, site, pct_cover_hard_coral_combined, depth_avg, aca_geomorphic, management_rules

Load Darling et al 2019 ('NEE') data: 
```{r}
dat_nee <- read_csv(here("globalprep","analysis", "coral-counterfactual","data_dl",
                         "Darling_etal_NEE","coral-model-data.csv")) %>% 
  clean_names()

dat_nee <- dat_nee %>% mutate(cc_id = seq(1:nrow(dat_nee)), 
         cc_id = str_c("nee_",cc_id))

dat_nee

# 2584 surveys; 2010-2016
dat_nee %>% tabyl(management) %>% adorn_totals()
dat_nee$year %>% hist() # 2010-2016
dat_nee %>% tabyl(country)

```

Select variables
```{r}

dat_nee1 <- dat_nee %>% 
  select(country, site, longitude, latitude, 
         sumcover, year, depth_m, habitat, management,
         method, cc_id)

# rename to combine with mermaid: 
dat_nee1 <- dat_nee1 %>% 
  rename(depth_avg = depth_m, 
         pct_hardcoral = sumcover) %>% 
  mutate(db = "darling_etal_2019",
         mgmt_simple = case_when(management == "No-take" ~ "no take",
                                 management == "Fished" ~ "open access",
                                 management == "Restricted" ~ "restricted")) %>% 
  select(-management)

# Convert all FJ 'open access' to 'restricted'
dat_nee1 <- dat_nee1 %>% 
    mutate(mgmt_simple = 
           case_when((country == "Fiji" & mgmt_simple == "open access") ~ "restricted",
                     TRUE ~ mgmt_simple))

dat_nee1 %>% tabyl(country,mgmt_simple)

# Check for dupes
# Only 2 duplicate sites (Singapore; US); lat/lons, methods, and mgmt differ. leave for now.
dat_nee1 %>% get_dupes(country,site)

dat_nee1 %>% skim()
dat_nee1 %>% tabyl(mgmt_simple) %>% adorn_totals() # 2584
dat_nee1 %>% tabyl(db)
```

## 1c. UQ photo quadrat data

```{r}
source(here("globalprep","analysis","coral-counterfactual", "01-format-uq.R"))
dat_uq

dat_uq1 <- dat_uq %>% 
  mutate(site = str_c("uq",surveyid, sep = "_"))

dat_uq1
```

## 1d. ETP data
Adds: 32 sites, 49 observations

```{r}
source(here("globalprep","analysis","coral-counterfactual", "01-format-etp.R"))
# use dat_etp0 instead of (averaged) dat_etp

dat_etp1 <- dat_etp0 %>% 
  mutate(db = "romero_torres_etal_2020",
         method = "unknown") %>% 
  rename(pct_hardcoral = coral_cover,
         latitude = lat, 
         longitude = lon,
         year = survey_year) %>% 
  select(db, method, country, site, latitude, longitude, year, pct_hardcoral)

dat_etp1 %>% tabyl(year) %>% 
  adorn_totals()

dat_etp1

```


# 2. Create "ccmod_df" with combined MERMAID, Darling et al NEE, UQ, and ETP data

```{r}
# 5422 surveys 
ccmod_df <- dat_mer1 %>% 
  bind_rows(dat_nee1) %>% 
  bind_rows(dat_uq1) %>%
  bind_rows(dat_etp1) %>% 
  arrange(country, site, year) %>% 
  select(db, country, site, latitude, longitude, depth_avg,
         year, method, pct_hardcoral, mgmt_simple)

ccmod_df <- ccmod_df %>% 
  mutate(ccmod_id = seq(1:nrow(ccmod_df)), 
         ccmod_id = str_c("ccmod_",ccmod_id), 
         method = str_to_lower(method), 
         method_cat = case_when(method == "bleaching" ~ "pitlitbleach",
                                method == "lit" ~ "pitlitbleach", 
                                method == "pit" ~ "pitlitbleach",
                                method == "unknown" ~ "pitlitbleach",
                                method == "photo" ~ "photo",
                                TRUE ~ NA_character_))

# Table S1 - # of surveys
ccmod_df %>% tabyl(db, method) %>% adorn_totals() #%>% view()
ccmod_df %>% tabyl(db, method_cat) %>% adorn_totals() #%>% view()

# Table S2 - survey years (2010 - 2021)
ccmod_df %>% tabyl(year) #%>% view()

# CHECK THESE. These are from combined MERMAID surveys; need to figure out best way to account for these differences.
ccmod_df %>% tabyl(mgmt_simple, db)

# Check: all FIJI sites listed as "open access" should be "restricted" 
ccmod_df %>% tabyl(country, mgmt_simple) %>% filter(country == "Fiji")
  

ccmod_df %>% skim() # 5422 surveys

ggplot(ccmod_df, aes(x = pct_hardcoral, fill = db))+
  geom_histogram()+
  facet_wrap(~db, scales = "free_y")

#ggsave(here("globalprep","analysis","coral-counterfactual","plots", "pct_hardcoral_dist_db.png"), width = 12, height = 8)


```


# 3. Average %CC by grid cell

## 3a. Load LRP cells
```{r}

# load data info
allreefs_info <- read_excel(here("globalprep", "analysis", "coral-counterfactual","data_dl",
          "wcs-local-reef-pressures","key.xlsx"))

allreefs_info

# load dataset (original allreefs.Rdata transformed using "01-format-allreefs.R")

load(here("globalprep", "analysis",
          "coral-counterfactual","data_dl",
          "wcs-local-reef-pressures", "allreefs_WGS84.RData"))

allreefs
```

# 4. Spatial join & summarise

```{r}

ccmod_sf <- ccmod_df %>% 
  st_as_sf(coords = c("longitude","latitude"),
           crs = 4326)

# spatial join coral cover data & lrp
# 1672 cells with coral cover data (of 54596 coral reef cells) -- 3.06% of reef cells
# 712 cells have 1 survey associated; 959 cells have > 1 survey
ccmod_5km_df <- st_join(ccmod_sf, allreefs) %>% 
  as_tibble() %>% 
  drop_na(OBJECTID) %>% 
  group_by(OBJECTID) %>% 
  summarise(pct_hardcoral = median(pct_hardcoral, na.rm =T),
            pct_hardcoral_min = min(pct_hardcoral, na.rm = T),
            pct_hardcoral_max = max(pct_hardcoral, na.rm = T),
            pct_hardcoral_nsurveys = n(),
            depth_avg = median(depth_avg, na.rm = T),
            sample_yr_median = median(year, na.rm = T),
            sample_yr_min = min(year, na.rm = T),
            sample_yr_max = max(year, na.rm = T),
            mgmt_simple_comb = paste(unique(mgmt_simple[!is.na(mgmt_simple)]), 
                                collapse = "; "),
            method_comb = paste(unique(method_cat[!is.na(method_cat)]), 
                                collapse = "; "),
            method_all  = paste(method_cat, collapse = "; ")) %>% 
  ungroup() %>% 
  mutate(mgmt_highest = case_when(str_detect(mgmt_simple_comb, "no take") ~ "no take", 
                                  str_detect(mgmt_simple_comb, "restricted") ~ "restricted",
                                  str_detect(mgmt_simple_comb, "open access") ~ "open access",
                                             TRUE ~ NA_character_)) %>% 
  select(-mgmt_simple_comb, -method_all) %>% 
  mutate(method_cat = if_else(str_detect(method_comb, ";"), "mixed", method_comb))

ccmod_5km_df %>% skim()
ccmod_5km_df %>% tabyl(mgmt_highest)
ccmod_5km_df %>% tabyl(method_comb)
ccmod_5km_df %>% tabyl(method_cat)
ccmod_5km_df %>% tabyl(year_avg)
ccmod_5km_df %>% filter(pct_hardcoral_nsurveys == 1) %>% nrow()
ccmod_5km_df %>% filter(pct_hardcoral_nsurveys > 1) %>% nrow()

ccmod_5km_df %>% filter(pct_hardcoral_nsurveys > 1,
                        sample_yr_min != sample_yr_max) %>% view()



# MGMT_SIMPLE: CHECK AND UPDATE MERMAID SITES! HERE! ALSO: ADD 'UNKNOWN' and 'restricted' for sites with mixed management categories.

lrp_cc <- allreefs %>% 
  as_tibble() %>% 
  left_join(ccmod_5km_df, by = "OBJECTID")

lrp_cc_sf <- 
  lrp_cc %>% 
  st_as_sf(crs = 4326)

```

# 5. Save things

## Save combined site-based data
```{r}
# Save table
ccmod_df %>% 
  write_rds(here("globalprep","analysis","coral-counterfactual","outputs",
             "cc_surveys_combn.RDS"))

# Save geojson
fn = here("globalprep","analysis","coral-counterfactual", "outputs", 
          "cc_surveys_combn.geojson")

if(file.exists(fn)){
  file.remove(fn)
}

st_write(ccmod_sf, dsn = fn, driver = "GeoJSON", append = F,layer = "cc_surveys")

```

## Save coral cover data aligned with 5km LRP grids
```{r}
# RDS
lrp_cc_sf %>% 
  write_rds(here("globalprep","analysis","coral-counterfactual",
                              "outputs","allreefs_coralcover.RDS"))

# geojson
fn = here("globalprep","analysis","coral-counterfactual", "outputs", 
          "allreefs_coralcover.geojson")

if(file.exists(fn)){
  file.remove(fn)
}

st_write(lrp_cc_sf, dsn = fn, driver = "GeoJSON", append = F)
```

## LRP Centroids
Save LRP (allreefs) centroids as *.csv
- Used in a lot of the 01-format scripts

```{r}
allreefs_centroids <- allreefs %>% 
  select(OBJECTID) %>% 
  st_centroid()

allreefs_centroids

allreefs_coords <- st_coordinates(allreefs_centroids$geometry) %>% 
  bind_cols(OBJECTID = allreefs_centroids$OBJECTID) %>% 
  rename(longitude = X, 
         latitude = Y) %>% 
  select(OBJECTID, longitude, latitude)

allreefs_coords %>% 
  write_csv(here("globalprep","analysis","coral-counterfactual",
          "outputs","allreefs_centroids.csv"))

# centroids for LRP cells with coral cover data
allreefs_cc_centroids <- lrp_cc_sf %>% 
  filter(!is.na(pct_hardcoral)) %>% 
  select(OBJECTID) %>% 
  st_centroid()


allreefs_cc_coords <- st_coordinates(allreefs_cc_centroids$geometry) %>% 
  bind_cols(OBJECTID = allreefs_cc_centroids$OBJECTID) %>% 
  rename(longitude = X, 
         latitude = Y) %>% 
  select(OBJECTID, longitude, latitude)

allreefs_cc_coords %>% 
  write_csv(here("globalprep","analysis","coral-counterfactual",
          "outputs","allreefs_cc_centroids.csv"))
  
```

## Outputs for Julia

- Lat / Long, database, cell ID for the ~5,400 coral cover sites we have assembled so far
- Example code for aligning to LRP grids (in R, library 'sf')
- Leaflet examples

```{r}
# sites: 
cc_sites <- ccmod_df %>% 
  select(db, country, site, latitude, longitude) %>% 
  distinct()

cc_sites %>% write_csv(here("globalprep","data-requests","julia",
                            "cc_sites.csv"))
```

